### 2017-11-15 Draft2/ConvertRawArray.Blurb.tex
Work

Users often want to convert dynamic arrays to raw arrays. There are multiple reasons for this: 1) Raw arrays hold on to exactly $n$ memory to store $n$ elements. However, dynamic and growth arrays typically use more than $n$ memory, so they need not grow every time an item is appended. 2) Sometimes, functions in third-party code only accept raw arrays.

For growth arrays, the case is even more compelling: 3) Since they are not contiguous as a whole, they have worse data locality (even if the number of discontinuities is $O(\log n)$). 4) Raw arrays can be indexed much faster than growth arrays.

>>>>

Users often want to convert dynamic arrays to raw arrays. There are multiple reasons for this: 1) Raw arrays hold on to exactly $n$ memory to store $n$ elements. However, dynamic and growth arrays typically use more than $n$ memory, so that they need not grow every time an item is appended. 2) Often, functions in third-party code only accept raw arrays.

For growth arrays, the case is even more compelling: 3) Since they are not contiguous as a whole, they have worse data locality (although the number of discontinuities is only $O(\log n)$). 4) Raw arrays can be indexed much faster than growth arrays.

### 2017-11-15 Draft2/Iterating.Growth.tex
Work

The iteration algorithm for growth arrays is not quite as simple. As shown in the previous section, growth arrays' random access algorithm is expensive; thus, it is best avoided. The following algorithm implements iteration without using the random access algorithm.

\begin{algorithm}[H]
	\caption{``$\TextFor \LclItem \TextIn \VarList$'' algorithm \TextGrowthArray}
	\begin{algorithmic}
		\For{$\LclBuffer \TextIn \VarList.\FieldTail$}
			\For{$\LclItem \TextIn \LclBuffer$}
				\State $\TextDoSomethingWith \LclItem$
			\EndFor
		\EndFor
		\For{$\LclIndex = 0,\ \LclIndex < \VarList.\FieldHeadSize,\ \LclIndex \gets \LclIndex + 1$}
			\State $\LclItem \gets \VarList.\FieldHead[\LclIndex]$
			\State $\TextDoSomethingWith \LclItem$
		\EndFor
	\end{algorithmic}
\end{algorithm}

>>>>

The iteration algorithm for growth arrays is not quite as simple. As shown in Section \ref{subsec:Indexing}, growth arrays' random access algorithm can be expensive; thus, it is best avoided. The following algorithm implements iteration without using the random access algorithm.

\begin{algorithm}[H]
	\caption{``$\TextFor \LclItem \TextIn \VarList$'' algorithm \TextGrowthArray}
	\begin{algorithmic}
		\For{$\LclBuffer \TextIn \VarList.\FieldTail$}
			\For{$\LclItem \TextIn \LclBuffer$}
				\State $\TextDoSomethingWith \LclItem$
			\EndFor
		\EndFor
		\For{$\LclIndex = 0,\ \LclIndex < \VarList.\FieldHeadSize,\ \LclIndex \gets \LclIndex + 1$}
			\State $\TextDoSomethingWith \VarList.\FieldHead[\LclIndex]$
		\EndFor
	\end{algorithmic}
\end{algorithm}

Since this algorithm loops over $n$ items, its time complexity is $O(n)$. There is nothing to suggest that this algorithm should be substantially slower than dynamic arrays' algorithm. Even though the inner loop is occasionally `interrupted' by having to switch buffers, the number of such interruptions is $O(\log n)$, so their cost is minuscule. Thus, the only substantial work done by this algorithm is iterating the buffers, which is roughly the same as the work done by dynamic arrays' algorithm.
